---
title: 'Foucaultize: Graphics step-by-step'
author: "krissacrates"
date: "6/2/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Wordcloud

First install packages needed:

```{r packages}
## Install 'wordcloud2' package if needed
packs <- installed.packages()
packs.match <- match("wordcloud2", packs)  
if (is.na(packs.match)) require("wordcloud2")
library(wordcloud2)     # load packagnoune
```

Have your lemmatized files ready. Use `foucaultize()` function to prepare the files
if needed. It is assumed that data is ready in the `./output-files` folder in your
working directory.

```{r chooseTable, eval=FALSE}
## Cannot be evaluated by Knit due to interactive menu() command
books <- c("History of Madness", "The Birth of the Clinic", "The Order of Things", 
           "The Archaeology of Knowledge", "Other - I will provide the acronym")
acronyms <- c("HF", "NC", "MC", "AS")
myBook <- switch (menu(books, title = "Choose the book:") + 1,
                NA, case1 = acronyms[1],
                    case2 = acronyms[2],
                    case3 = acronyms[3],
                    case4 = acronyms[4],
                    case5 = "other")    # interactive items cannot be knitted into HTML
if (is.na(myBook)) warning("No book, no fun!")
if (myBook == "other") myBook <- readline(prompt = "What is your book's acronym? (eg. HF): ")
```

Now load and adjust the dataset:

```{r wordcloud2}
## I set 'myBook' manually because previous code chunk cannot be knitted in HTML.
myBook <- "HF"  # Remove if you use the previous code chunk
############  
pathFile <- paste0("./output-files/", myBook, "_lemmas_nouns.csv")
cloud2.df <- read.csv(pathFile)
cloud2.df <- cloud2.df[, c("lemma", "n")]    # subset dataframe to 2 columns as expected
set.seed(8732)
wordcloud2(cloud2.df, color = "random-light", backgroundColor="black")  
```

Credit: Some other folks did a great job [here](https://www.r-bloggers.com/the-wordcloud2-library/).
  

## Frequency barchart

Another library needed:

```{r lattice}
packs.match <- match("lattice", packs)  
if (is.na(packs.match)) require("lattice")
library(lattice)     # load package
```

Then it's so easy: 

```{r barchart}
myFile <- paste0(myBook, "_lemmas_nouns.csv")
barchart(lemma ~ n, cloud2.df[1:15, ], col = "grey", xlab = "Term frequency", 
         main = paste("Lemma frequency in", myFile))
```


## Heatmap

Heatmap should represent distribution of occurences of the keyword throughout the book.

Firstly, we have to prepare data in the right format to plot a heatmap.

```{r heatmapData}
fullPath <- paste0("./output-files/", myBook, "_lemmas_full.csv")   # path to full data CSV
full.df <- read.csv(fullPath)   # read the file
# Subset nouns only
a <- which(full.df$wclass == "noun")    # vector of row numbers with nouns
nouns.vector <- full.df[a, "lemma"] # subset only nouns based on 'a' vector
remove(full.df) # cleanup memory from a big file

# Some variable ad-hod definitions
keyword <- "experience" # set keyword for which we plot the heatmap of occurences
clusters <- 20L     # number of clusters (chunks) to split the whole book into

logical.vector <- grepl(keyword, nouns.vector, ignore.case = TRUE)  # each occurence make TRUE otherwise FALSE
volume.clusters <- ceiling( length(logical.vector) / clusters )  # number of keyword items in a cluster
chunks <- split(logical.vector, ceiling(seq_along(logical.vector)/volume.clusters)) # create a list of chunks (= clusters)
sum.clusters <- sapply(chunks, sum) # calculate sum of TRUE values each cluster - creates a numeric vector of sums
heat.matrix <- matrix(sum.clusters, nrow = 1, ncol = length(sum.clusters) )  
str(heat.matrix)
```

To create a heatmap, instal `plotly` package. Inspiration source: [plot.ly/r/heatmaps/](plot.ly/r/heatmaps/)

```{r plotly}
packs.match <- match("plotly", packs)  
if (is.na(packs.match)) require("plotly")
library(plotly)     # load packagnoune
```

Now, we are ready to plot a heatmap.

```{r heatmap1}
#Optimize heatmap visuals
vals <- unique(scales::rescale(c(heat.matrix)))
o <- order(vals, decreasing = FALSE)
cols <- scales::col_numeric("Blues", domain = NULL)(vals)
colz <- setNames(data.frame(vals[o], cols[o]), NULL)

# Create a heatmap
p <- plot_ly(y = keyword, z = heat.matrix, colorscale = colz, type = "heatmap", height = 200)
p
```



