---
title: "keywords_intersection"
author: "krissacrates"
date: "10 6 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

## Get data


```{r packages, eval=TRUE}
# evaluation is turned off in this code chunk
# "ggplot2" package needs to be active
library(ggplot2)
library(dplyr)
```

Find common keywords for HF and NC.

```{r getTables}

HF_top100 <- read.csv("./output-files/HF_lemmas_nouns.csv")
x <- HF_top100[1:20, c("lemma", "freq")]
HF_top100 <- HF_top100[1:20, "lemma"]
HF_top100 <- as.character(HF_top100)

NC_top100 <- read.csv("./output-files/NC_lemmas_nouns.csv")
y <- NC_top100[1:20, c("lemma", "freq")]
NC_top100 <- NC_top100[1:20, "lemma"]
NC_top100 <- as.character(NC_top100)


HF_NC_intersection <- intersect(HF_top100, NC_top100)
intersection.df <- data.frame(lemma = HF_NC_intersection)

intersection.df <- merge(intersection.df, x, by = "lemma")
intersection.df <- plyr::rename(intersection.df, c("freq" = "HF_freq"))

intersection.df <- merge(intersection.df, y, by = "lemma")
intersection.df <- plyr::rename(intersection.df, c("freq" = "NC_freq"))

cumul_freq <- intersection.df$HF_freq + intersection.df$NC_freq
intersection.df <- data.frame(intersection.df, cumulative_freq = cumul_freq)
intersection.df <- intersection.df[order(intersection.df[, "cumulative_freq"], decreasing = TRUE), ]
intersection.df
```


Make graph of the intersected keywords. Standard barchart:

```{r graph1}
# source: http://www.r-graph-gallery.com/80-circular-barplot/
# make data
data = data.frame(lemma = intersection.df$lemma, cumulative_freq = intersection.df$cumulative_freq )
# Usual bar plot :
graph1 <- ggplot(data, aes(x = lemma, y = cumulative_freq ,fill = cumulative_freq )) + geom_bar(width = 0.85, stat="identity", position = position_stack(reverse = TRUE)) + coord_flip()
graph1
```


Circular barchart:

```{r graph2}
graph2 <- ggplot(data, aes(x = lemma, y = cumulative_freq, fill = cumulative_freq )) + geom_bar(width = 0.85, stat="identity") + 

# To use a polar plot and not a basic barplot
coord_polar(theta = "y") +

#Remove useless labels of axis
xlab("") + ylab("") +

#Increase ylim to avoid having a complete circle
ylim(c(0,0.008)) + 

#Add group labels close to the bars :
geom_text(data = data, hjust = 1.5, size = 5, aes(x = lemma, y = 0, label = lemma)) +

#Remove useless legend, y axis ticks and y axis text
theme(legend.position = "none" , axis.text.y = element_blank() , axis.ticks = element_blank())
```


Add another intersection with MC:

```{r table2}
MC_top100 <- read.csv("./output-files/MC_lemmas_nouns.csv")
z <- MC_top100[1:20, c("lemma", "freq")]
MC_top100 <- MC_top100[1:20, "lemma"]
MC_top100 <- as.character(MC_top100)

HF_NC_MC_intersection <- intersect(HF_NC_intersection, MC_top100)
# Intersection for first 20 keywords in MC with previous set.
HF_NC_MC_intersection
```

There are only 4 keywords. There are only 2 with conceptual power: order and knowledge.
Trully, we can consider it as conceptual foundation for archaeology. But it is still very narrow. 
Let's try first 100 keywords from MC to intersect with the previous set.

```{r table3}
MC_top100 <- read.csv("./output-files/MC_lemmas_nouns.csv")
z <- MC_top100[1:100, c("lemma", "freq")]
MC_top100 <- MC_top100[1:100, "lemma"]
MC_top100 <- as.character(MC_top100)

HF_NC_MC_intersection <- intersect(HF_NC_intersection, MC_top100)
HF_NC_MC_intersection
```

Great! In the second foundation layer, we have 3 more: truth, experience, and body.
Let's create a new table for intersection of previous set with MC:

```{r getTable2}
intersection2.df <- data.frame(lemma = HF_NC_MC_intersection)

intersection2.df <- merge(intersection2.df, x, by = "lemma")
intersection2.df <- plyr::rename(intersection2.df, c("freq" = "HF_freq"))

intersection2.df <- merge(intersection2.df, y, by = "lemma")
intersection2.df <- plyr::rename(intersection2.df, c("freq" = "NC_freq"))

intersection2.df <- merge(intersection2.df, z, by = "lemma")
intersection2.df <- plyr::rename(intersection2.df, c("freq" = "MC_freq"))

cumul_freq <- intersection2.df$HF_freq + intersection2.df$NC_freq + intersection2.df$MC_freq
intersection2.df <- data.frame(intersection2.df, cumulative_freq = cumul_freq)
intersection2.df <- intersection2.df[order(intersection2.df[, "cumulative_freq"], decreasing = TRUE), ]
intersection2.df

```

Let's visualize:

```{r graph3}
# make data
data = data.frame(lemma = intersection2.df$lemma, cumulative_freq = intersection2.df$cumulative_freq )
# Usual bar plot :
graph3 <- ggplot(data, aes(x = lemma, y = cumulative_freq ,fill = cumulative_freq )) + geom_bar(width = 0.85, stat="identity", position = position_stack(reverse = TRUE)) + coord_flip()
graph3
```


